---
title: Seasonal Bump Functions
date: 2019-04-11
author: Dr. Juan Camilo Orduz
categories: Machine Learning
tags: r,  machine_learning
slug: bump_func
---

Motivated by the nice talk on [Winning with Simple, even Linear, Models](https://www.youtube.com/watch?v=68ABAU_V8qI) by Vincent D. Warmerdam, I briefly describe how to construct certain class of [bump functions](https://en.wikipedia.org/wiki/Bump_function) to encode seasonal variables in [R](https://www.r-project.org/).

```{r, echo=FALSE}
library(knitr)
knitr::opts_knit$set(warning=FALSE, message=FALSE)
```

# Prepare Notebook

```{r, warning=FALSE, message=FALSE}
library(glue)
library(lubridate)
library(magrittr)
library(tidyverse)
```

# Generate Data 

Let us generate a time sequence variable stored in a tibble. 

```{r}
# Define time sequence. 
t <- seq.Date(from = as.Date("2017-07-01"), to = as.Date("2019-04-01"), by = "day")
# Store it in a tibble. 
df <- tibble(t = t)
```

# Symmetric Bump Function

We want to generate a gaussian-like curve around a specific date. The width of the bump function is parametrized by a non-negative number $\varepsilon \geq 0$. For $\varepsilon = 0$ we generate a dummy variable. Namely, for $\varepsilon \neq 0$, 

$$
f_{x_0, \varepsilon}(x) = \exp\left(- \left(\frac{x - x_0}{\varepsilon}\right)^2 \right)
$$

In R, 

```{r}
bump_func <- function(x, x_0 = 0, epsilon = 1) {
  # If `epsilon` = 0, we encode it as a dummy variable.
  if (epsilon == 0) {
    as.numeric(x == x_0)
  } else {
    delta <-  as.numeric(x - x_0)
    exp(- (delta / epsilon)^2)
  }
}
```

## Examples

Let us asumme we want to model the effect of Christmas `2018-12-24`. 

- $\varepsilon = 0$ will generate a dummy variable. 

```{r, fig.align="center"}
epsilon <- 0

df %>% 
  mutate(y = bump_func(x = t, 
                       x_0 = as.Date("2018-12-24"), 
                       epsilon = epsilon)) %>% 
  ggplot() +
  geom_line(mapping = aes(x = t, y = y)) +
  xlab("Date") +
  ylab("") +
  ggtitle(label = glue("Bump Seasonal Variable - epsilon = {epsilon}"))

```

- $\varepsilon = 8$ will generate a bump function around `2018-12-24`. 

```{r, fig.align="center"}
epsilon <- 8

df %>% 
  mutate(y = bump_func(x = t, 
                       x_0 = as.Date("2018-12-24"),
                       epsilon = 8)) %>% 
  ggplot() +
  geom_line(mapping = aes(x = t, y = y)) +
  xlab("Date") +
  ylab("") +
  ggtitle(label = glue("Bump Seasonal Variable - epsilon = {epsilon}"))

```

Let us see how to do it for repeated dates and events:

```{r, fig.align="center"}
epsilon <- 8

christmas_dates <- c(as.Date("2017-12-24"), as.Date("2018-12-24"))
halloween_dates <- c(as.Date("2017-10-31"), as.Date("2018-10-31"))

df %>% 
  mutate(
    y_christmas = map(
      .x = christmas_dates, 
      .f = ~ bump_func(x = t, x_0 = .x, epsilon = epsilon)) %>% 
      reduce(.f = ~ .x + .y),
    
    y_halloween = map(
      .x = halloween_dates, 
      .f = ~ bump_func(x = t, x_0 = .x, epsilon = epsilon)) %>% 
      reduce(.f = ~ .x + .y)) %>% 
  gather(key = Event, value = value, ... = y_christmas, y_halloween) %>% 
  # Plot variables.
  ggplot() +
  geom_line(mapping = aes(x = t, y = value, color = Event)) +
  xlab("Date") +
  ylab("") +
  ggtitle(label = glue("Bump Seasonal Variable - epsilon = {epsilon}"))
```

# Asymmetric Bump Function

Sometimes the effect before and after the seasonal variable is note the same. To model this we can tweak the bump function defined above so that: 

  - It is not symmetric.
  - It can have a drop after the seasonal variable (an potentially have different effect sizes). 

A natural candidate to fulfill these requirements is 

$$
a_{-} f_{x_0, \varepsilon_{-}}(x)I_{\{x \leq x_0\}} \pm a_{+} f_{x_0, \varepsilon_{+}}(x) I_{\{x > x_0\}}
$$
where:

  - $I_{\{x \leq x_0\}}$ and $I_{\{x > x_0\}}$ denote the corresponing [indicator functions](https://en.wikipedia.org/wiki/Indicator_function).

  - $f_{x_0, \varepsilon_{\pm}}$ denotes the bump function defined above. 
  
  - $\varepsilon_{-}, \varepsilon_{+} >0$ denote the width (variance) before and after $x_0$ respectively. 
  
  - $a_{-}, a_{+} >0$ denote the amplitudes before and after $x_0$ respectively.  

We now write this function in R. 

```{r, fig.align="center"}
asymmestric_bump_func <- function(
  # Input vector.
  x,
  # Maximum of bump function.
  x_0, 
  # Width of bump function before `x_0`.
  epsilon_minus = 1, 
  # Width of bump function after `x_0`.
  epsilon_plus = 1, 
  # Amplitude of bump function before `x_0`.
  a_minus = 1, 
  # Amplitude of bump function after `x_0`.
  a_plus = 1, 
  # Wether to have a drom after `x_0`.
  drop = FALSE
  ) {
  
  delta <- as.numeric(x - x_0)
  # Define indicator functions.
  indicator_minus <- as.numeric(delta < 0)
  indicator_plus <- as.numeric(delta >= 0)
  
  # Calculate bump function components. 
  y_minus <- a_minus * indicator_minus * bump_func(x = x, x_0 = x_0, epsilon = epsilon_minus) 
  y_plus <- a_plus*indicator_plus * bump_func(x = x, x_0 = x_0, epsilon = epsilon_plus)
  # Calculate sign depending on the drop choice. 
  drop_sign <- 2*as.numeric(drop) - 1
  
  y <- y_minus  - drop_sign*y_plus
  
  return(y)
}
```

## Examples

Let us ilustrate the results for Christmas `2017-12-24`.

### Without Drop 

```{r, fig.align="center"}
df  %>% 
  mutate(y =  asymmestric_bump_func(x = t,
                                    x_0 = as.Date("2017-12-24"),
                                    epsilon_minus = 20,
                                    epsilon_plus = 4, 
                                    drop = FALSE)) %>% 
  # Filter for visualization (zoom in). 
  filter(t < "2018-02-01") %>% 
  ggplot() +
  geom_line(mapping = aes(x = t, y = y)) +
  xlab("Date") +
  ylab("") +
  ggtitle(label = glue("Asymmetric Bump Function"))
```

### With Drop 

```{r, fig.align="center"}
df  %>% 
  mutate(y =  asymmestric_bump_func(x = t,
                                    x_0 = as.Date("2017-12-24"),
                                    epsilon_minus = 15,
                                    epsilon_plus = 4, 
                                    a_minus = 1, 
                                    a_plus = 0.3,
                                    drop = TRUE)) %>% 
  # Filter for visualization (zoom in). 
  filter(t < "2018-02-01") %>% 
  ggplot() +
  geom_line(mapping = aes(x = t, y = y)) +
  xlab("Date") +
  ylab("") +
  ggtitle(label = glue("Asymmetric Bumb Function with Drop"))
```


