In this blog post I want to describe the explicit computation of the Laplacian on differential forms on the $2$-Torus $T^2\subset \mathbb{R}^3$. This surface can be obtained by rotating, for $0<r<a$, the circle $(x-a)^2+y^2=r^2$ around the $z$-axis. This surface can be parametrized (locally) by the equations
$$
x = (a+r\cos u)\cos v,\\
y = (a+r\cos u)\sin v,\\
z = r\sin u,
$$ 

where $0<u,v<2\pi$. 

```{r}
a <- 4
r <- 1

u_seq <- seq(from = 0, to = 2*pi, by = 0.1)
v_seq <- seq(from = 0, to = 2*pi, by = 0.1)

local_parametrization <- function(a, r, u, v) {
  
  x <- (a + r*cos(u))*cos(v)
  y <- (a + r*cos(u))*sin(v)
  z <- r*sin(u)
  
  local_chart <- list(x = x, y = y, z = z)
  local_chart_df <- data.frame(local_chart)
  
  return(local_chart_df)
}
```


```{r}
library(plotly)

#plot_ly(data = local_chart_df, x = ~ x, y = ~ y, z = ~ z)
```


## Euclidean Metric 

The metric on $T^2$ induced by the Euclidean metric of $\mathbb{R}^3$ is given by
$$
ds^2=r^2du^2+(a+r\cos u)^2dv^2
$$

## Volume Form

and the associated volume element is $\text{vol}_{T^2}=r(a+r\cos u)du\wedge dv$.

## Hodge Operator 

First we compute the action of the Hodge star operator. As we did before we just use the definition, for example to compute $*du$ we compute
$$
du\wedge * du =\langle du, du \rangle \text{vol}_{T^2}=r^{-2}r(a+r\cos u)du\wedge dv,
$$
so we see that $*du=r^{-1}(a+r\cos u)dv$. Similarly we obtain $*dv=-r(a+r\cos u)^{-1}dv$.\\

Now we want to compute the Hodge-Dirac operator for an $S^1$-invariant form on $\BB{T} $ which can always be written as
$$
\omega =f_0(u)+f_1(u)du+f_2(u)dv+f_3(u)du\wedge dv.
$$
Its exterior derivative is 
$$
d\omega  = f'_0(u)du+f'_2(u)du\wedge dv.
$$

## Dirac Operator

## Laplacian 