
In this notebook post I want to describe the explicit computation of the Laplacian on differential forms on the $2$-Torus $T^2\subset \mathbb{R}^3$. This surface can be obtained by rotating the circle $(x-a)^2+y^2=r^2$ around the $z$-axis ($0<r<a$). Locally, this surface can be parametrized by the equations
$$
x = (a+r\cos u)\cos v,\\
y = (a+r\cos u)\sin v,\\
z = r\sin u,
$$ 

where $0<u,v<2\pi$. 

**Remark:** Note that, topologically, $T^2$ can be identified with the product of two circles. In particular, it is a [Lie group](https://en.wikipedia.org/wiki/Lie_group).

## Visualization

Let us use this local parametrization to visualize the 2-torus for $a=4$ and $r=1$. 

```{r}
a <- 4
r <- 1

# Parameter defining how many points to take. 
frac_density <- 0.03
# Get parameter grid. 
u_seq <- seq(from = 0, to = 2*pi, by = frac_density)
v_seq <- seq(from = 0, to = 2*pi, by = frac_density)

# Generate surface from local coordinates. 
local_parametrization <- function(a, r, u, v) {
  
  x <- (a + r*cos(u))*cos(v)
  y <- (a + r*cos(u))*sin(v)
  z <- r*sin(u)
  
  local_chart <- list(x = x, y = y, z = z)
  local_chart_df <- data.frame(local_chart)
  
  return(local_chart_df)
}

coordinates_df <- expand.grid(... = u_seq, v_seq)
colnames(coordinates_df) <- c("u", "v")

local_parametrization_list <- apply(
  X = coordinates_df, 
  MARGIN = 1, 
  FUN = function(x) {local_parametrization(a = a, r = r, u = x[1] , v = x[2])}
)

local_parametrization_df <- do.call(what = rbind, args = local_parametrization_list)
rownames(local_parametrization_df) <- c(1:nrow(local_parametrization_df))
```

Now we use a 3d-scatter plot:

```{r, fig.align="center", warning=FALSE, message=FALSE}
library(plotly)

p <- plot_ly(
  data = local_parametrization_df,
  x = ~ x, 
  y = ~ y, 
  z = ~ z, 
  color = ~ z
)

layout(p = p, scene = list(zaxis = list(range = c(-4,4))))
```


## Euclidean Metric 

The metric on $T^2$ induced by the Euclidean metric of $\mathbb{R}^3$ is given by
$$
g^{T^2}=r^2du^2+(a+r\cos u)^2dv^2
$$
Let us prove this. Given the local parametrization $\phi(u, v) = (x(u, v), y(u, v), z(u, v))$ defined above, we need to compute $\partial_u\phi$ and $\partial_v\phi$. The components of the induced metric are then 
$$
g^{T^2} = E_{u}du^2 + E_{uv}du\otimes dv + E_{v}dv^2,
$$
where 
$$
E_{u} = \langle\partial_{u}\phi, \partial_{u}\phi\rangle_{\mathbb{R}^3}, \\
E_{uv} = \langle\partial_{u}\phi, \partial_{v}\phi\rangle_{\mathbb{R}^3}, \\
E_{v} = \langle\partial_{v}\phi, \partial_{v}\phi\rangle_{\mathbb{R}^3}
$$
We now calculate:
$$
\frac{\partial x(u, v)}{\partial u} = -r\sin u\cos v, \quad \frac{\partial y(u, v)}{\partial u} = -r\sin u\sin v  \quad \text{and} \quad \frac{\partial z(u, v)}{\partial u} = r \cos u, 
$$
which implies:
$$
E_u = ( -r\sin u\cos v)^2 + (-r\sin u\sin v )^2 + (r \cos u)^2 = r^2
$$
Similarly, 
$$
\frac{\partial x(u, v)}{\partial v} = - (a+r\cos u)\sin v , \quad \frac{\partial y(u, v)}{\partial v} = (a+r\cos u)\cos v   \quad \text{and} \quad \frac{\partial z(u, v)}{\partial v} = 0, 
$$
and 
$$
E_v = (- (a+r\cos u)\sin v)^2 + (- (a+r\cos u)\cos v)^2 + 0^2 = (a+r\cos u)^2
$$
Finally, we see that the crossed term vanishes: 
$$
E_{uv} = ( -r\sin u\cos v)(- (a+r\cos u)\sin v ) + (-r\sin u\sin v )(a+r\cos u)\cos v = 0
$$

**Remark:** In matrix form, with respect to the basis $du, dv$, we can write the metric as 
$$
g^{T^2}
=
\left(
\begin{array}{cc}
r^2 & 0 \\
0 & (a+r\cos u)^2
\end{array}
\right)
$$

## Volume Form

The volume form $\text{vol}_{T^2}$ on the 2-torus is defined by the relation $\text{vol}_{T^2} = \sqrt{\det(g^{T^2})} du\wedge dv$ (the sign $du\wedge dv$ is a choise orientation). Hence,  $\text{vol}_{T^2}=r(a+r\cos u)du\wedge dv$.

## Hodge Star Operator 

Recall that the [Hodge star operator](https://en.wikipedia.org/wiki/Hodge_star_operator) is definied by the relation
$$
\alpha \wedge *\beta = \langle \alpha, \beta \rangle \text{vol}_{T^2}, 
$$
for $\alpha, \beta \in \bigwedge^*T^*(T^2)$ (or differential forms). The inner product is defined on the space of $n$-forms. In particular, 
$$
\langle du, du \rangle = \frac{1}{r^2} \quad\text{and}\quad \langle dv, dv \rangle = \frac{1}{(a+r\cos u)^2}.
$$
Note that $*1 = \text{vol}_{T^2}$. 

Let us now compute the Hodge star operator on the $1$-forms $du$ and $dv$:

$$
\langle du, du \rangle \text{vol}_{T^2}=r^{-2}r(a+r\cos u)du\wedge dv,
$$
so we see that $*du=r^{-1}(a+r\cos u)dv$. Similarly we obtain $*dv=-r(a+r\cos u)^{-1}du$.

## Dirac Operator

The Dirac operator of the Clifford bundle of differential forms is given by $D = d + d^{\dagger}$, where $d^\dagger = -*d*$ (as the dimension is even). For example, for a smooth function $f\in C^{\infty}(T^2)$, we have
$$
D f = df = \frac{\partial f}{\partial u}du + \frac{\partial f}{\partial v}dv
$$
and for a $1$-form $f du$ we obtain
$$
D(fdu) = d(fdu) -f*d*du = - \frac{\partial f}{\partial dv} du\wedge dv-f*d\left(r^{-1}(a+r\cos u)dv\right)
$$
Hence, 
$$
D(fdu) = - \frac{\partial f}{\partial dv} du\wedge dv + f \sin u * (du\wedge dv) = - \frac{\partial f}{\partial dv} du\wedge dv + \frac{f \sin u}{r(a + r \cos u)} 
$$

## Laplacian 

The Laplacian is defined by $\Delta = D^2$. Let us compute it explicitly for a smooth function $f\in C^{\infty}(T^2)$. As we have seen $Df = df$, hence
$$
\Delta = D(df) = (d + d^\dagger) df = d^\dagger df
$$
On the one hand side, 
$$
df = \frac{\partial f}{\partial u}du + \frac{\partial f}{\partial v}dv  
$$

Let us study the first term:
$$
d^{\dagger} \left(\frac{\partial f}{\partial u}du\right) = -*d*\left(\frac{\partial f}{\partial u}du\right) \\
= - * d\left(r^{-1}(a+r\cos u)\frac{\partial f}{\partial u} dv\right) \\
-*\left(-\sin u\frac{\partial f}{\partial u} + r^{-1}(a+r\cos u)\frac{\partial^2 f}{\partial u^2}\right)du\wedge dv\\
= \frac{1}{r(a + r\cos u)}\left(\sin u\frac{\partial f}{\partial u} - r^{-1}(a+r\cos u)\frac{\partial^2 f}{\partial u^2}\right)
$$
The second term can be computed similarly
$$
d^{\dagger} \left(\frac{\partial f}{\partial v}dv\right) = -*d*\left(\frac{\partial f}{\partial v}dv\right) \\
=  * d\left(r(a+r\cos u)^{-1}\frac{\partial f}{\partial v} du\right) \\
= - * \left(r(a+r\cos u)^{-1}\frac{\partial^2 f}{\partial v^2} du\wedge dv\right) \\
= - \frac{1}{r(a + r\cos u)}\left(r(a+r\cos u)^{-1}\frac{\partial^2 f}{\partial v^2}\right) 
$$
Hence, 
$$
\Delta f =\frac{1}{r(a + r\cos u)}\left(\sin u\frac{\partial f}{\partial u} - r^{-1}(a+r\cos u)\frac{\partial^2 f}{\partial u^2}\right) - \frac{1}{r(a + r\cos u)}\left(r(a+r\cos u)^{-1}\frac{\partial^2 f}{\partial v^2}\right) 
$$
which simplifies to 
$$
\Delta f = -\frac{1}{r^2}\frac{\partial^2 f}{\partial u^2} - \frac{1}{(a + r\cos u)^2}\frac{\partial^2 f}{\partial v^2} + \frac{\sin u}{r(a + r\cos u)}\frac{\partial f}{\partial u}.
$$ 

For one and two forms the computation is very similar.  
