---
title: The volume of the $d$-ball via Monte Carlo Simulation
date: 2020-02-25
author: Dr. Juan Camilo Orduz
categories: geometry
tags: r,  geometry
slug: vol_d_ball
---

```{r, echo=FALSE}
library(knitr)
knitr::opts_knit$set(warning=FALSE, message=FALSE)
```

In this notebook we run Monte Carlo simulations to estimate the volume of the $d$-ball $B^{d}:=\{x \in \mathbb{R}^d : ||x|| \leq 1\}$. There are many proofs to get the analytic solution, see for example this [Wikipedia article](https://en.wikipedia.org/wiki/N-sphere). 

```{r}
library(tidyverse)
```

## Main Idea

Consider a square $A_{d}\subset \mathbb{R}$ centered at the origin with side length $2$. We estimate the volume of the $d$-ball $B^{d}:=\{x \in \mathbb{R}^d : ||x|| \leq 1\}\subset A^{d}$ by sampling uniformly from $A$ and computing the proportions of vectors having length less or equal than one. Note that $\text{vol}(A^{d}) = 2^{d}$. 

```{r}
area_square <- function (d) { 2^d }
```


Let us write a function to generate such samples:

```{r}
# d: dimension.
# n: number os samples.
get_samples_df <- function (d, n) {
   map(.x = 1:d, 
      .f = function(i) { 
        samples <- runif(n = n, min = -1, max = 1)
        tibble(!!str_c("d_", as.character(i)) := samples)
  }
  ) %>% reduce(.f = ~ bind_cols(.x, .y))
}
```

Let us see an example for $n=2$. 

```{r, fig.align="center"}
# Get samples. 
samples_df <- get_samples_df(d = 2, n = 1E4)
# Compute L2-norm.
norm_2 <- samples_df %>% mutate_all(.funs = ~ .x^2) %>% rowSums() 
# Inside the ball condtion.
is_in_ball <- (norm_2 <= 1)
samples_df <- samples_df %>% mutate(is_in_ball = is_in_ball) 

# plot.
samples_df %>% 
  ggplot(mapping = aes(x = d_1, y = d_2, color = is_in_ball)) +
  geom_point(alpha = 0.4) +
  scale_color_brewer(palette = "Set1") +
  theme(aspect.ratio = 1)
```

Let us compute the area ratio:

```{r}
area_square(d = 2)*(sum(samples_df$is_in_ball) / nrow(samples_df))
```

Recall that the true area of the $2$-ball is $\pi$, so the estimation is quite close. 

## Volume Computation 

Let us now generalize the above computation for arbitrary dimensions. 

```{r}
compute_vol_ball <- function (d, n) {
  
  samples_df <- get_samples_df(d = d, n = n)

  norm_d <- samples_df %>% mutate_all(.funs = ~ .x^2) %>% rowSums()

  is_in_ball <- norm_d < 1

  vol <- 2^(d)*(sum(is_in_ball) / n)
  
  return(vol)
}
```


## Simulation 

Next, we write a function to run the simulation. We do this as a function of the dimension $1\leq d \leq d_{max}$. 

```{r}
run_simulation <- function (d_max, n, N) {
  
  simulation_list <- map(.x = 1:d_max, .f = function (d) {
    map_dbl(.x = 1:N, .f = ~ compute_vol_ball(d = d, n = n))
  })
  
  simulation_df <- tibble(volume = simulation_list) %>% 
    rowid_to_column(var = "d") %>% 
    unnest(cols = volume) 
  
  return(simulation_df)
}
```


```{r}
# Simulation parameters. 
d_max <- 25
n <- 1E5
N <- 100

# Run simulation. 
simulation_df <- run_simulation(d_max = d_max, n = n, N = N)
# Get simulation mean.
mean_simulation_df <- simulation_df %>% 
  group_by(d) %>% 
  summarise(volume = mean(volume))
```

The analytic expression for the volume is^[[Wikipedia: N-Sphere](https://en.wikipedia.org/wiki/N-sphere)]

$$
\text{vol}(B^{d}) = \frac{\pi^{d/2}}{\Gamma(d/2 + 1)}
$$
where $\Gamma$ denotes the special [gamma function](https://en.wikipedia.org/wiki/Gamma_function). 

```{r}
vol_d_ball <- function (d) {
  pi^{(d)/2} / gamma(d/2 + 1)
}

vol_df <- tibble(
  d = 1:d_max , 
  volume_true = map_dbl(.x = 1:d_max, .f = vol_d_ball)
)
```

## Results

Let us compare the simulation results with the analytic solution: 

```{r, fig.align="center"}
set1_pal <- RColorBrewer::brewer.pal(n = 9, name = "Set1")

ggplot() +
  geom_point(data= simulation_df, mapping = aes(x = d, y = volume), color = set1_pal[2], alpha = 0.3) +
  geom_point(data = mean_simulation_df, mapping = aes(x = d, y = volume, color="sample")) +
  geom_line(data = mean_simulation_df, mapping = aes(x = d, y = volume, color="sample"), alpha = 0.5) +
  geom_point(data = vol_df, mapping = aes(x = d, y = volume_true, color = "true")) +
  geom_line(data = vol_df, mapping = aes(x = d, y = volume_true, color = "true"), alpha = 0.5) +
  labs(title = "Volume d-Balls",subtitle = "Monte Carlo Simulation", xlabel = "dimension") +
  scale_color_manual(values=c(set1_pal[2], set1_pal[1])) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

```{r, fig.align="center"}
ggplot() +
  geom_point(data = mean_simulation_df, mapping = aes(x = d, y = volume, color="sample")) +
  geom_line(data = mean_simulation_df, mapping = aes(x = d, y = volume, color="sample"), alpha = 0.5) +
  geom_point(data = vol_df, mapping = aes(x = d, y = volume_true, color = "true")) +
  geom_line(data = vol_df, mapping = aes(x = d, y = volume_true, color = "true"), alpha = 0.5) +
  labs(title = "Volume d-Balls",subtitle = "Monte Carlo Simulation (Mean) vs Analytic Expression", xlabel = "dimension") +
  scale_color_manual(values=c(set1_pal[2], set1_pal[1])) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

It is very interesting to see that the ball in dimension $d=5$ has the biggest volume and then the volume tends to zero. But why? Does this make sense? Well, this plot is actually misleading as we are comparing apples with oranges as we are comparing different units. That is, we are comparing meters ($m$) with square meters ($m^2$) and so on.

We can compare the ratio of the volumes of each ball and the corresponding squares, i.e. $\text{vol}(B^2)/\text{vol}(A^d)$ which does not have any units. 

```{r, fig.align="center"}
vol_df %>% 
  mutate(vol_prop = volume_true / 2^(d + 1)) %>% 
  ggplot(mapping = aes(x = d, y = vol_prop)) +
  geom_point() +
  geom_line(alpha = 0.5) +
  labs(title = "Ratio d-Ball/d-Square", y="")
```

Note that this ratio is always decreasing and converges to zero. This is related to the [Curse of dimensionality](https://en.wikipedia.org/wiki/Curse_of_dimensionality). See [Exploring the Curse of Dimensionality - Part I.](https://juanitorduz.github.io/exploring-the-curse-of-dimensionality-part-i./) and [Exploring the Curse of Dimensionality - Part II.](https://juanitorduz.github.io/exploring-the-curse-of-dimensionality-part-ii./)

**Remark** If we define the *diameter* of a set $S\subset \mathbb{R}^{d}$as 
$$
\text{diam}(S):= \sup_{x, y\in S}||x - y||
$$
Then $\text{diam}(B^{d})=2$, where the $\sup$ is attained when considering two antipodal points, and $\text{diam}(A^{d})=2\sqrt{d}$, where again the $\sup$ by considering the points $(-1, -1, \cdots, -1)$ and $(1, 1, \cdots, 1)$. 